### 2022-03-09 misc/solutions/prob-method-sols.tex
Added two new questions that I did on break

\def\given{\,|\,}
\def\biggiven{\,\big|\,}
\def\F{{\cal F}}

\maketitle{The Probabilistic Method}{exercise solutions by}{Marcel K. Goh}{17 February 2022}

@@

>>>>

\def\given{\,|\,}
\def\biggiven{\,\big|\,}
\def\A{{\cal A}}
\def\B{{\cal B}}
\def\C{{\cal C}}
\def\F{{\cal F}}
\def\G{{\cal G}}

\maketitle{The Probabilistic Method}{exercise solutions by}{Marcel K. Goh}{9 March 2022}

@@

\section 6. Correlation Inequalities

\proclaim Exercise 6.1. Let $G$ be a graph and let $P$ denote the probability that a random subgraph
of $G$ obtained by picking each edge of $G$ with probability $1/2$ independently is connected (and spanning).
Let $Q$ denote the probability that in a random two-colouring of $G$, where each edge is chosen,
randomly and independently, to be either red or blue, the red graph and the blue graph are both connected
(and spanning). Is $Q\le P^2$?

\solution The answer is yes. Let $\A$ be the family of subsets $A\subseteq E$ such that $G' = (V,A)$
is connected. Let $\C$ be the family of subsets $C\subseteq E$ such that $G'' = (V, E\setminus C)$ is
connected. Adding any edge to $A\in \A$ keeps the set in $\A$, and removing any edge from $C\in \C$ keeps
the set in $\C$, so $\A$ is monotone increasing and $\C$ is monotone decreasing. Note that if $S$ is a
random subset of edges obtained by taking each edge independently with probability $1/2$, then
$\pr\{S = A\} = \pr\{S = E\setminus A\}$ for any fixed $A\subseteq E$, so we find that
$ P = \pr\{\A\} = \pr\{\C\}$. On the other hand, $Q$ is the probability that both $S$ and $E\setminus S$
induce connected subgraphs; i.e., $Q = \pr\{S\in \A\cap \C\}$. Applying Proposition~6.3.1 now gives
$$Q = \pr\{ \A\cap \C\} \le \pr\{\A\} \pr\{\C\} = P^2,$$
as desired.\slug

\proclaim Exercise 6.2. A family of subsets $\G$ is called {\it intersecting} if $G_1\cap G_2\ne \emptyset$
for all $G_1,G_2\in \G$. Let $\F_1, \F_2, \ldots, \F_k$ be $k$ intersecting families of subsets of $X = [n]$.
Prove that
$$\biggl|\bigcup_{i=1}^k \F_i\biggr| \le 2^n - 2^{n-k}.$$

\proof We do this by induction on $k$. If $k=1$, then the upper bound of $2^{n-1}$ on the size of an
intersecting family $\F$ is clear from the fact that $\F$ may contain at most one element of each pair
$(A,X\setminus A)$, and there are $2^{n-1}$ such pairs.
For the induction step, we split
$$\bigcup_{i=1}^k \F_i = \bigcup_{i=1}^{k-1} \F_i \cup \F_k;$$
call the first term on the right-hand side $\A$ and let $\C = 2^X\setminus \F_k$, so that
$$\bigcup_{i=1}^{k-1} \F_i \cup \F_k = \A \cup \F_k = (\A \cap \C) \cup \F_k.$$
Since we are trying to form an upper bound on this quantity,
we might as well assume that $\A$ is maximal; i.e., adding any set to $\F_i$ for $1\le i< k$ causes
$\F_i$ to no longer be intersecting. The claim is that $\A$ is monotone increasing. To see this, suppose
that $A\supseteq B$ and $B\in \A$. This means that $B\in \F_i$ for some $1\le i<k$, and so for any $C\in \F_i$,
$A\cap C \supseteq B\cap C\ne \emptyset$, meaning that $A\in \F_i\subseteq \A$ as well, otherwise we could
add $A$ to $\A$ and contradict maximality. We have actually shown
that the union of any $m$ maximal intersecting families is monotone increasing, so in particular $\F_k$ is
as well. But this means that $\C = 2^X\setminus \F_k$ is monotone decreasing, since if $C\notin \F_k$
and $B\subseteq C$ with $B\in \F_k$, then $\F_k$ would not be monotone increasing. We are now in a position
to apply the cardinality version of Proposition~6.3.1. It gives
$$\biggl|\bigcup_{i=1}^k \F_i\biggr| = |\A\cap \C| + |\F_k|
\le {|\A|\cdot|\C|\over 2^n} + |\F_k|
\le \biggl(1-{1\over 2^{k-1}}\biggr) |\C| + |\F_k|,$$
where in the last inequality we used the induction hypothesis $|\A| \le 2^n - 2^{n-k+1}$. Since
$|\C| + |\F_k| = 2^n$ and the coefficient of the latter is larger than the coefficient of the former
in the above, the entire expression is largest when $|\F_k| = |\C| = 2^{n-1}$. In this case, we have
$$\biggl|\bigcup_{i=1}^k \F_i\biggr| \le \biggl(1-{1\over 2^{k-1}}\biggr) 2^{n-1} + 2^{n-1} = 2^n-2^{n-k},$$
completing the induction and the proof.\slug
